---
title: "GridLMMnet"
output: html_notebook
---

```{r}
library(GridLMM)
library(glmnet)
```

This demonstrates the use of the `GridLMMnet` function.

Create some simulated data:
```{r}
nG = 300
nR = 3
n = nG*nR
data = data.frame(Group = factor(rep(1:nG,each = nR)),ID = 1:n)
p = 1000
X = matrix(sample(c(0,1),n*p,replace=T),n,p)
rownames(X) = data$ID


s2_g = 0.8
s2_e = 0.2
b1 = 10
b2 = p-b1
beta = c(rep(.3,b1),rep(0,b2))

y = X %*% beta + sqrt(s2_g)*rnorm(nG)[data$Group] + sqrt(s2_e)*rnorm(n)

library(lme4)
m1 = lmer(y~X[,1:b1]+(1|Group),data)
v1 = as.data.frame(VarCorr(m1))$vcov
v1[1]/sum(v1)
v1
sum(v1)

```

Use glmnet to get the lasso solutions without random effects
```{r}
res0 = glmnet(X,y,alpha=1)
```

Standard coefficient-by-lambda plot. Red = true covariates, black= false covariates.
```{r}
plot(res0,'lambda',col = c(rep(2,b1),rep(1,b2)))
```

Run GridLMMnet

```{r}
res1 = GridLMMnet(y~1+(1|Group),data = data,X = X,h2_step = 0.025)
```

Standard coefficient-by-lambda plot. Red = true covariates, black= false covariates.
```{r}
plot(res1,'lambda',col = c(rep(2,b1),rep(1,b2)))
```

Plot the estimate variance component h^2
```{r}
plot(res1$h2s~ log(res1$lambda))
abline(h=s2_g/(s2_g+s2_e))
```

Plot estimate of s2_g
```{r}
plot(res1$h2s*res1$s2s~ - log(res1$lambda),ylim = c(0,var(y)))
abline(h=s2_g)
abline(h=v1[1],col=2)
```

plot estiamte of s2_e
```{r}
plot((1-res1$h2s)*res1$s2s~ log(res1$lambda),ylim = c(0,var(y)))
abline(h=s2_e)  # simulated data
abline(h=v1[2],col=2) # true estimate
```


plot estimate of s2 = s2_g + s2_e
```{r}
plot(res1$s2s~ log(res1$lambda),ylim = c(0,var(y)))
abline(h=s2_g+s2_e)  # simulated data
abline(h=sum(v1),col=2) # true estimate
```





Compare to ggmix
```{r}
library(ggmix)
K = tcrossprod(model.matrix(~0+factor(Group),data))
sK = svd(K)
res2 = ggmix(X,y,U = sK$u[,1:nG],D = sK$d[1:nG])
# res2 = ggmix(X,y,kinship = K,n_nonzero_eigenvalues = nG)
```


Standard coefficient-by-lambda plot. Red = true covariates, black= false covariates.
```{r}
plot(res2,col = c(rep(2,b1),rep(1,b2)))
```


```{r}
plot(res2$eta[1,]~ log(res2$lambda))
```

